---
title: "Homework 1 - Time Series Visualization"
author: "Prof. Dr. Stephan Trahasch"
date: 'Submission date: 17.12.2020'
output:
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp3)
library(fpp2) 
```

# Exercise 1
Use the help menu to explore what the series `gold`, `woolyrnq` and `gas` represent. These are available in the `forecast` package.

```{r, echo=TRUE, cache=TRUE, eval=FALSE}
help(gold)
help(woolyrnq)
help(gas)
# to do

```

Use `autoplot` to plot each of these in separate plots.

```{r, echo=TRUE, cache=TRUE}
# to do

autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)
```

**What is the frequency of each commodity series?**
Hint: apply the `frequency()` function.

```{r, echo=TRUE, cache=TRUE}
# to do
help("frequency")
frequency(gold)
frequency(woolyrnq)
frequency(gas)
```

Use `which.max()` to spot the outlier in the `gold` series. 
**Which observation was it?**

```{r, echo=TRUE, cache=TRUE}
# to do

help("which.max")
which.max(gold)
which.max(woolyrnq)
which.max(gas)

```

# Exercise 2 

Import the file `tute1.csv` from the folder data and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
tute1 <- read.csv("data/tute1.csv", header = TRUE)
View(tute1)
```

We convert the data to time series. The `[,-1]` removes the first column which contains the quarters as we don't need them now.

```{r, echo=TRUE, cache=TRUE}
mytimeseries <- ts(tute1[, -1], start = 1981, frequency = 4)
```

Construct time series plots of each of the three series and use 'facets = TRUE' .

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(mytimeseries, facets=TRUE)
```

Check what happens when you don't include `facets=TRUE`.

```{r, echo=TRUE, cache=TRUE, dependson="exmyts"}
# to do
autoplot(mytimeseries, facets=FALSE)
```

# Exercise 3

Import some monthly Australian retail data from the foler data. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
retaildata <- readxl::read_excel("data/retail.xlsx", skip = 1)
```

The second argument (`skip=1`) is required because the Excel sheet has two header rows.

Select one of the time series as follows (but replace the column name with your own chosen column):

```{r, echo=TRUE, cache=TRUE}
myts <- ts(retaildata[, "A3349627V"], frequency = 12, start = c(1982, 4))
```

Explore your chosen retail time series using the following functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf`

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(myts)
ggseasonplot(myts)
ggsubseriesplot(myts)
gglagplot(myts,lags=4)
ggAcf(myts)
```

**Can you spot any seasonality, cyclicity and trend? What do you learn about the series?**

Your answer:
Trend: Positiv wachsender Trend
Zyklisch: Jedes Jahr wiederkehrend 
Sesonalität: Schwierig aus den Grafiken zu erkennen

# Exercise 4
Create time plots of the following time series: goog, writing, fancy, a10, h02.

    + Use `help()` to find out about the data in each series.
    + For the `goog` plot, modify the axis labels and title.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# to do
autoplot(goog)
autoplot(writing)
autoplot(fancy)
autoplot(a10)
autoplot(h02)
#gglagplot(goog)
#ggAcf(goog)

#autoplot(writing)
#ggseasonplot(writing)
#ggsubseriesplot(writing)
#gglagplot(writing)
#ggAcf(writing)
```

Use the `ggseasonplot` and `ggsubseriesplot` functions to explore the seasonal patterns in the following time series:  `writing`, `fancy`, `a10`, `h02`.

- What can you say about the seasonal patterns?
- Can you identify any unusual years?


```{r, echo=TRUE}
# writing
# to do
ggseasonplot(writing)
ggsubseriesplot(writing)
```

Your answer:
Es gibt eine Sesonale Komponente, die immer im August des Jahres statt findet.Ich konnte kein ausergewöhnliches Jahr feststellen.

a10:Jan
h02:Jan

```{r, echo=TRUE}
# fancy
# to do

help("fancy")
ggseasonplot(fancy)
ggsubseriesplot(fancy)

```

Your answer:#
Die Sesonale Komponente liegt immer im November und Dezember mit einem Anstieg.
Die Jahre 1992 und 1993 haben einen besonderst starken Anstieg im Nov. und Dez.

```{r, echo=TRUE}
# a10
# to do
help(a10)
ggseasonplot(a10)
ggsubseriesplot(a10)
```

Your answer:
Januar ist der Extremwert als Sesonalekomponente, die wiederkehrend ist.
Von 1991 -  2000 verlaufen die Linen noch eher sehr flach 
Danach sehr gezackt bzw. sehr unregelmäßig

```{r, echo=TRUE}
# h02
# to do
help(h02)
ggseasonplot(h02)
ggsubseriesplot(h02)
```

Your answer:
Januar Maximum und Februar Minimum als Sesionale Kompnente, die dann auch über Februar bis zum Januar stätig ansteigt. 

# Exercise 5
Use the following graphics functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf` and explore features from the following time series: `hsales`, `usdeaths`, `bricksq`, `sunspotarea`, `gasoline`.

    + Can you spot any seasonality, cyclicity and trend?
    + What do you learn about the series?

Example for hsales:

```{r, echo=TRUE, cache=TRUE}
autoplot(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
gglagplot(hsales)
ggAcf(hsales)
```

  + Seasonality evident in all plots
  + Cyclicity seen in first two plots
  + No trend
  + ACF only shows seasonality. Cycle length too long to show up here.

```{r, echo=TRUE, cache=TRUE}
# to do
help(usdeaths)
autoplot(usdeaths)
ggseasonplot(usdeaths)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths)

plot(stl(usdeaths,"periodic"),main="Accidental deaths in USA")
```

Your answer:
seasonality: Immer im July ist ein Extremwert über die Jahre zu verseichnen
cyclicity: keine Aussage, konnte ich nicht 
and trend: Der Trend über die Jahre bis Mitte des Jahres 1976 abgeflacht und steigt nun wiederan.

# Exercise 6

The `pigs` data shows the monthly total number of pigs slaughtered in Victoria, Australia, from Jan 1980 to Aug 1995. Use `mypigs <- window(pigs, start=1990)` to select the data starting from 1990. Use `autoplot` and `ggAcf` for `mypigs` series and compare these to white noise plots.

```{r, echo=TRUE, cache=TRUE}
# White noise
#set.seed(1711)
#y <- ts(rnorm(50))
#autoplot(y) + ggtitle("White noise")

mypigs <- window(pigs, start=1990)
autoplot(mypigs)
ggAcf(mypigs)
```

```{r}
ggAcf(mypigs)
```


```{r, echo=TRUE, cache=TRUE}
# todo
set.seed(mypigs)
y <- ts(rnorm(mypigs))
autoplot(y) + ggtitle("White noise")

# White noise
```

Your answer:


# Exercise 7

`dj` contains 292 consecutive trading days of the Dow Jones Index. Use `ddj <- diff(dj)` to compute the daily changes in the index. Plot `ddj` and its ACF. Do the changes in the Dow Jones Index look like white noise?

```{r}
# todo
ddj <- diff(dj)
plot(ddj)
ggAcf(ddj)

set.seed(ddj)
y <- ts(rnorm(ddj))
autoplot(y) + ggtitle("White noise")
```

Your answer:
Es ist White noise vorhanden.
